Accelerating structure-function mapping using the ViVa webtool to mine natural variation

### 2019-04-02 manuscript/04-methods.Rmd
revisions done, back to coauthors

### Gene and Transcript Accession Information
Information on the genes and transcripts including chromosomal coordinates, start and end location, and transcript length were pulled from Araport11 [@cheng_araport11_2017]. The TAIR10 database, found at http://arabidopsis.org, was accessed via the `biomart` function, using the biomaRt R-package. The Araport11 full genome general feature format file, which can also be found on the TAIR website (https://www.arabidopsis.org/download/index-auto.jsp?dir=%2Fdownload_files%2FGenes%2FAraport11_genome_release), has been embedded in the R-package as GRanges object, `gr`. Gene identifiers used in this study are in Table \@ref(tab:gene-ids-table).
```{r gene-ids-table}
AFBs <- read.csv(system.file("shiny-app", "gene_family_data",
                      "AFB_gene_ids.csv", package = "r1001genomes"))
IAAs <- read.csv(system.file("shiny-app", "gene_family_data",
                      "IAA_gene_ids.csv", package = "r1001genomes"))
TPLs <- read.csv(system.file("shiny-app", "gene_family_data",
                      "TPL_gene_ids.csv", package = "r1001genomes"))
ARFs <- read.csv(system.file("shiny-app", "gene_family_data",
                      "ARF_gene_ids.csv", package = "r1001genomes"))
gene_ids <- bind_rows(AFBs, IAAs, TPLs, ARFs)
kable(gene_ids, col.names = c("Arabidopsis AGI locus identifier", "Gene symbol", "Clade/class"),caption = "Full list of genes used in this study, by identifier, symbol and classification.")
```

@@

>>>>

### Gene and Transcript Accession Information
Information on the genes and transcripts including chromosomal coordinates, start and end location, and transcript length were downloaded from Araport11 as a General Feature Format (GFF) file [@cheng_araport11_2017]. The Araport11 full genome general feature format file, which can also be found on the TAIR website (https://www.arabidopsis.org/download/index-auto.jsp?dir=%2Fdownload_files%2FGenes%2FAraport11_genome_release), has been embedded in the R-package as GRanges object, `gr`. The TAIR10 database, found at http://arabidopsis.org, was accessed via the `biomart` function, using the biomaRt R-package. Gene identifiers used in this study are in Table \@ref(tab:gene-ids-table).
```{r gene-ids-table}
AFBs <- read.csv(system.file("shiny-app", "gene_family_data",
                      "AFB_gene_ids.csv", package = "r1001genomes"))
IAAs <- read.csv(system.file("shiny-app", "gene_family_data",
                      "IAA_gene_ids.csv", package = "r1001genomes"))
TPLs <- read.csv(system.file("shiny-app", "gene_family_data",
                      "TPL_gene_ids.csv", package = "r1001genomes"))
ARFs <- read.csv(system.file("shiny-app", "gene_family_data",
                      "ARF_gene_ids.csv", package = "r1001genomes"))
gene_ids <- bind_rows(AFBs, IAAs, TPLs, ARFs)
kable(gene_ids, col.names = c("Arabidopsis AGI locus identifier", "Gene symbol", "Clade/class"),caption = "Full list of genes used in this study, by identifier, symbol and classification.")
```

@@

## Testing
A group consisting of undergraduate and graduate students, postdoctoral and postbaccalaureate researchers (the authors of this publication) were assembled to test the functionality of ViVa. Testers were asked to use the ViVa web interface to analyze the genetic variation within a gene family or clade within a family of the auxin nuclear signaling pathway. Testers were provided with a brief vignette on how to use ViVa to formulate new hypotheses or support existing hypotheses about gene function and evolution, which has been expanded to the "An Overview of ViVa" section. The testers had background knowledge of these genes as members of laboratories studying auxin nuclear signaling. Testers met weekly with the developers to discuss their user experience and their findings. Tester experiences, issues, and suggestions were incorporated into the ViVa software. The results collected by these analyses are summarized in the "Visualizing Variation within the auxin signaling pathway" section and [Supplemental Data](#si).

### 2019-04-02 manuscript/06-supplement.Rmd
revisions done, back to coauthors

```{r AFB_ alignment}
alignment <- alignCDS(gene_info$transcript_ID)
# make an alignment data frame
aln_df <- makeAlnDF(alignment[[2]])

@@

p <-ggplot(aln_df, aes(x = aln_pos, y = tair_symbol)) +
  geom_raster(mapping = aes(fill = strength,
                            alpha = gap)) +
  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
                     labels = c("aligned", "gap"), 
                     guide = guide_legend(override.aes = 
                                            list(fill = c("grey85", "white")),
                       title = "alignment",
                       direction = "vertical")) + 
   scale_fill_viridis(breaks = range(SNPeff_order$strength),
                      limits = range(SNPeff_order$strength),
                      na.value = "grey85",
                      direction = -1,
                      labels = c("weak", "strong"),
                      option = "A", 
                      guide = guide_colorbar(ticks = FALSE, 
                                             title.vjust = .8, 
                                             title.position = "top", 
                                             title.hjust = 0.5)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
  labs(x = "codon position in alignment", y = "", fill = "variant functional effect") +
  theme_logo(base_family = "Helvetica") +
  theme(panel.grid = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        #panel.background = element_rect(fill = "grey85", color = "white"),
        legend.position = c("bottom"),
        legend.key = element_rect(linetype = "solid"),
        legend.spacing = unit(4, "char"), 
        legend.box.spacing = unit(0.1,"line")) 
p

@@

(ref:full-AFBs-alignment) **Alignment of the AFB family** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects. Red indicates missense variants. Color scale is explained in [Methods](#methods).

@@

The AFB4 and AFB5 receptors have an N-terminal extension prior to their F-box domains. This extension had very high nonsynonymous diversity (Figure \@ref(fig:full-AFBs-alignment), suggesting that this extension does not play an important functional role in these proteins. Additionally, two frameshift variants and one stop-gained variant were observed in *AFB4* supporting its pseudogenization.

For simplicity we have included only the alignment of the class A Aux/IAAs in the main manuscript. For completion we include here the complete alignment of the family.

(ref:full-IAA-alignment) **Alignment of the complete Aux/IAA family.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects. Red indicates missense variants. Color scale is explained in [Methods](#methods).

```{r full-IAA-alignment, fig.cap="(ref:full-IAA-alignment)", fig.height=8}
IAAs <- alignCDS(gene_info$transcript_ID)
# make an alignment data frame
IAAs <- makeAlnDF(IAAs[[2]])
IAAs <- left_join(IAAs, dplyr::select(gene_info, "tair_locus",
                                       "tair_symbol", "transcript_ID"),
                        by = c("transcript_ID" = "transcript_ID"))
IAAs$tair_symbol %<>% as.factor()
IAAs$tair_symbol %<>% factor(levels = IAA_order, 
                                       ordered = TRUE)
IAAs$tair_symbol %<>% fct_rev()
VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
coding_VCF <- getCodingDiv(VCF)
IAAs <- addSNPsToAlnDF(aln_df = IAAs,SNPs = coding_VCF, effect_order = SNPeff_order)
IAAs$gap <- IAAs$seq_pos == "-"

p <-ggplot(IAAs, aes(x = aln_pos, y = tair_symbol)) +
  geom_raster(data = IAAs, 
             # mapping = aes(fill = effects)) +
              mapping = aes(fill = strength,
                            alpha = gap)) +
  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
                     labels = c("aligned", "gap"), 
                     guide = guide_legend(override.aes = 
                                            list(fill = c("grey85", "white")),
                       title = "alignment",
                       direction = "vertical")) + 
   scale_fill_viridis(breaks = range(SNPeff_order$strength),
                      limits = range(SNPeff_order$strength),
                      na.value = "grey85",
                      direction = -1,
                      labels = c("weak", "strong"),
                      option = "A", 
                      guide = guide_colorbar(ticks = FALSE, 
                                             title.vjust = .8, 
                                             title.position = "top", 
                                             title.hjust = 0.5)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
  labs(x = "codon position in alignment", y = "", fill = "variant functional effect") +
  theme_logo(base_family = "Helvetica") +
  theme(panel.grid = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        #panel.background = element_rect(fill = "grey85", color = "white"),
        legend.position = c("bottom"),
        legend.key = element_rect(linetype = "solid"),
        legend.spacing = unit(4, "char"), 
        legend.box.spacing = unit(0.1,"line")) 
p

```{r, IAA6-div-plot, fig.cap="(ref:IAA6-div-plot)"}
coding_variants <- VCF[VCF$Transcript_ID == gene_info[gene_info$tair_symbol == "IAA6", "transcript_ID"], ]
p <- plotCodingDiv(uniqueCodingVars = coding_variants)
p + annotate("rect", xmin = 106, xmax = 127, ymin = 0.001, ymax = 1, color = "lightgrey", alpha =.1) + annotate("text", x = 116.5, y = 0.5, label = "region of positive\nselection identified\nin Winkler et al.")
```
(ref:IAA6-div-plot) **IAA6 diversity plot.** Nucleotide diversity of variant positions throughout the IAA6 coding sequence are plotted and colored according to the effect of the variant alleles at each position. The region of positive selection identified by Winkler et al. is highlighed.

@@

```{r TPL-alignment-coding-vars}
alignment <- alignCDS(gene_info$transcript_ID)
#possible outgroups  AtLUG, At4g32551; AtLUH, At2g32700
# make an alignment data frame
aln_df <- makeAlnDF(alignment[[2]])

@@

(ref:full-TPLs-alignment) **Alignment of the TPL/TPR family.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects. Red indicates missense variants. Color scale is explained in [Methods](#methods).
```{r full-TPLs-alignment, fig.cap= "(ref:full-TPLs-alignment)"}
TPL_order <- fct_inorder(na.exclude(p$data$label[base::order(p$data$y, decreasing = TRUE)]), ordered = TRUE)

@@

p <-ggplot(aln_df, aes(x = aln_pos, y = tair_symbol)) +
  geom_raster(data = aln_df, 
             # mapping = aes(fill = effects)) +
              mapping = aes(fill = strength,
                            alpha = gap)) +
  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
                     labels = c("aligned", "gap"), 
                     guide = guide_legend(override.aes = 
                                            list(fill = c("grey85", "white")),
                       title = "alignment",
                       direction = "vertical")) + 
   scale_fill_viridis(breaks = range(SNPeff_order$strength),
                      limits = range(SNPeff_order$strength),
                      na.value = "grey85",
                      direction = -1,
                      labels = c("weak", "strong"),
                      option = "A", 
                      guide = guide_colorbar(ticks = FALSE, 
                                             title.vjust = .8, 
                                             title.position = "top", 
                                             title.hjust = 0.5)) +
  scale_y_discrete(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
  labs(x = "codon position in alignment", y = "", fill = "variant functional effect") +
  theme_logo(base_family = "Helvetica") +
  theme(panel.grid = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        #panel.background = element_rect(fill = "grey85", color = "white"),
        legend.position = c("bottom"),
        legend.key = element_rect(linetype = "solid"),
        legend.spacing = unit(4, "char"), 
        legend.box.spacing = unit(0.1,"line")) 
p

@@

```{r ARF-alignment-coding-vars}
alignment <- alignCDS(gene_info$transcript_ID)
# make an alignment data frame
aln_df <- makeAlnDF(alignment[[2]])

@@

(ref:full-ARFs-alignment) **Alignment of the full ARF family.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects. Red indicates missense variants. Color scale is explained in [Methods](#methods).

@@

>>>>

```{r AFB_ alignment}
alignment <- alignCDS(gene_info$transcript_ID)
consensus <- BALCONY::RealValET_conservativity(as.alignment(Biostrings::as.matrix(alignment[[2]])))
# make an alignment data frame
aln_df <- makeAlnDF(alignment[[2]])

@@

p <-ggplot(aln_df, aes(x = aln_pos, y = tair_symbol)) +
  geom_raster(mapping = aes(fill = strength,
                            alpha = gap)) +
  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
                     labels = c("aligned", "gap"), 
                     guide = guide_legend(override.aes = 
                                            list(fill = c("grey85", "white")),
                       title = "alignment",
                       direction = "vertical")) + 
   scale_fill_viridis(breaks = range(SNPeff_order$strength),
                      limits = range(SNPeff_order$strength),
                      na.value = "grey85",
                      direction = -1,
                      labels = c("weak", "strong"),
                      option = "A", 
                      guide = guide_colorbar(ticks = FALSE, 
                                             title.vjust = .8, 
                                             title.position = "top", 
                                             title.hjust = 0.5)) +
  scale_y_discrete(expand = expand_scale(add = c(1,2.5))) +
  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
  labs(x = "codon position in alignment", y = "", fill = "variant functional effect") +
  theme_logo(base_family = "Helvetica") +
  theme(panel.grid = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        #panel.background = element_rect(fill = "grey85", color = "white"),
        legend.position = c("bottom"),
        legend.key = element_rect(linetype = "solid"),
        legend.spacing = unit(4, "char"), 
        legend.box.spacing = unit(0.1,"line"))  +
  annotation_raster(matrix(gray(rescale(consensus, to = c(0,1))), nrow = 1),  xmin = -Inf, xmax = Inf, ymin = -10, ymax = 0.4)
p

@@

(ref:full-AFBs-alignment) **Alignment of the AFB family** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects. Red indicates missense variants. Color scale is explained in [Methods](#methods). In grayscale underneath the plot, alignment consensus is shown as measured by Evolutionary trace [@pluciennik_balcony_2018; @wilkins_evolutionary_2012], with high consensus positions in black and low consensus in white.

@@

```{r AFB1-map, fig.cap="(ref: AFB1-map)"}
knitr::include_graphics("app_figure/Map_figure.png")
```
(ref: AFB1-map) **Map of AFB1 Oligomerization domain variant accessions** (a) Map showing the two accessions with variants in the AFB1 oligomerization domain (b) The user selectable parameters of the mapping tab used to generate the map are provided as an example of using the SNP Mapping tab in the ViVa webtool. See Figure \@ref(fig:app-workflow) for details on filling in the parameters.

The AFB4 and AFB5 receptors have an N-terminal extension prior to their F-box domains. This extension had very high nonsynonymous diversity (Figure \@ref(fig:full-AFBs-alignment)), suggesting that this extension does not play an important functional role in these proteins. Additionally, two frameshift variants and one stop-gained variant were observed in *AFB4* supporting its pseudogenization.

@@

```{r TPL-alignment-coding-vars}
alignment <- alignCDS(gene_info$transcript_ID)
consensus <- BALCONY::RealValET_conservativity(as.alignment(Biostrings::as.matrix(alignment[[2]])))
#possible outgroups  AtLUG, At4g32551; AtLUH, At2g32700
# make an alignment data frame
aln_df <- makeAlnDF(alignment[[2]])

@@

(ref:full-TPLs-alignment) **Alignment of the TPL/TPR family.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects. Red indicates missense variants. Color scale is explained in [Methods](#methods). In grayscale underneath the plot, alignment consensus is shown as measured by Evolutionary trace [@pluciennik_balcony_2018; @wilkins_evolutionary_2012], with high consensus positions in black and low consensus in white.
```{r full-TPLs-alignment, fig.cap= "(ref:full-TPLs-alignment)"}
TPL_order <- fct_inorder(na.exclude(p$data$label[base::order(p$data$y, decreasing = TRUE)]), ordered = TRUE)

@@

p <-ggplot(aln_df, aes(x = aln_pos, y = tair_symbol)) +
  geom_raster(data = aln_df, 
             # mapping = aes(fill = effects)) +
              mapping = aes(fill = strength,
                            alpha = gap)) +
  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
                     labels = c("aligned", "gap"), 
                     guide = guide_legend(override.aes = 
                                            list(fill = c("grey85", "white")),
                       title = "alignment",
                       direction = "vertical")) + 
   scale_fill_viridis(breaks = range(SNPeff_order$strength),
                      limits = range(SNPeff_order$strength),
                      na.value = "grey85",
                      direction = -1,
                      labels = c("weak", "strong"),
                      option = "A", 
                      guide = guide_colorbar(ticks = FALSE, 
                                             title.vjust = .8, 
                                             title.position = "top", 
                                             title.hjust = 0.5)) +
  scale_y_discrete(expand = expand_scale(add = c(1,2.5))) +
  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
  labs(x = "codon position in alignment", y = "", fill = "variant functional effect") +
  theme_logo(base_family = "Helvetica") +
  theme(panel.grid = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        #panel.background = element_rect(fill = "grey85", color = "white"),
        legend.position = c("bottom"),
        legend.key = element_rect(linetype = "solid"),
        legend.spacing = unit(4, "char"), 
        legend.box.spacing = unit(0.1,"line")) +
  annotation_raster(matrix(gray(rescale(consensus, to = c(0,1))), nrow = 1),  xmin = -Inf, xmax = Inf, ymin = -10, ymax = 0.4)
p

@@

```{r ARF-alignment-coding-vars}
alignment <- alignCDS(gene_info$transcript_ID)
consensus <- BALCONY::RealValET_conservativity(as.alignment(Biostrings::as.matrix(alignment[[2]])))
# make an alignment data frame
aln_df <- makeAlnDF(alignment[[2]])

@@

(ref:full-ARFs-alignment) **Alignment of the full ARF family.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects. Red indicates missense variants. Color scale is explained in [Methods](#methods). In grayscale underneath the plot, alignment consensus is shown as measured by Evolutionary trace [@pluciennik_balcony_2018; @wilkins_evolutionary_2012], with high consensus positions in black and low consensus in white.

@@

For simplicity we have included only the alignment of the class A Aux/IAAs in the main manuscript. For completion we include here the complete alignment of the family (Figure \@ref(fig:full-IAA-alignment)).

(ref:full-IAA-alignment) **Alignment of the complete Aux/IAA family.** Protein sequences were aligned with DECIPHER [@wright_decipher_2015] and variants were mapped to this alignment and colored according to the predicted functional effect of the allele of strongest effect at that position, with light colors having weaker effects on function and darker colors stronger effects. Red indicates missense variants. Color scale is explained in [Methods](#methods). In grayscale underneath the plot, alignment consensus is shown as measured by Evolutionary trace [@pluciennik_balcony_2018; @wilkins_evolutionary_2012], with high consensus positions in black and low consensus in white.

```{r full-IAA-alignment, fig.cap="(ref:full-IAA-alignment)", fig.height=8}
IAAs <- alignCDS(gene_info$transcript_ID)
consensus <- BALCONY::RealValET_conservativity(as.alignment(Biostrings::as.matrix(IAAs[[2]])))
# make an alignment data frame
IAAs <- makeAlnDF(IAAs[[2]])
IAAs <- left_join(IAAs, dplyr::select(gene_info, "tair_locus",
                                       "tair_symbol", "transcript_ID"),
                        by = c("transcript_ID" = "transcript_ID"))
IAAs$tair_symbol %<>% as.factor()
IAAs$tair_symbol %<>% factor(levels = IAA_order, 
                                       ordered = TRUE)
IAAs$tair_symbol %<>% fct_rev()
VCF <- ldply(.data = VCF, .fun = subset, gt_GT != "0|0")
coding_VCF <- getCodingDiv(VCF)
IAAs <- addSNPsToAlnDF(aln_df = IAAs,SNPs = coding_VCF, effect_order = SNPeff_order)
IAAs$gap <- IAAs$seq_pos == "-"

p <-ggplot(IAAs, aes(x = aln_pos, y = tair_symbol)) +
  geom_raster(data = IAAs, 
             # mapping = aes(fill = effects)) +
              mapping = aes(fill = strength,
                            alpha = gap)) +
  scale_alpha_manual(values = c("TRUE" = 0, "FALSE" = 1), 
                     labels = c("aligned", "gap"), 
                     guide = guide_legend(override.aes = 
                                            list(fill = c("grey85", "white")),
                       title = "alignment",
                       direction = "vertical")) + 
   scale_fill_viridis(breaks = range(SNPeff_order$strength),
                      limits = range(SNPeff_order$strength),
                      na.value = "grey85",
                      direction = -1,
                      labels = c("weak", "strong"),
                      option = "A", 
                      guide = guide_colorbar(ticks = FALSE, 
                                             title.vjust = .8, 
                                             title.position = "top", 
                                             title.hjust = 0.5)) +
  scale_y_discrete(expand = expand_scale(add = c(1,2.5))) +
  scale_x_continuous(expand = c(0,0)) +  # expand increases distance from axis
  labs(x = "codon position in alignment", y = "", fill = "variant functional effect") +
  theme_logo(base_family = "Helvetica") +
  theme(panel.grid = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        #panel.background = element_rect(fill = "grey85", color = "white"),
        legend.position = c("bottom"),
        legend.key = element_rect(linetype = "solid"),
        legend.spacing = unit(4, "char"), 
        legend.box.spacing = unit(0.1,"line"))  +
  annotation_raster(matrix(gray(rescale(consensus, to = c(0,1))), nrow = 1),  xmin = -Inf, xmax = Inf, ymin = -10, ymax = 0.4)
p

```{r IAA6-div-plot, fig.cap="(ref:IAA6-div-plot)"}
coding_variants <- VCF[VCF$Transcript_ID == gene_info[gene_info$tair_symbol == "IAA6", "transcript_ID"], ]
p <- plotCodingDiv(uniqueCodingVars = coding_variants)
p + annotate("rect", xmin = 106, xmax = 127, ymin = 0.001, ymax = 1, color = "lightgrey", alpha =.1) + annotate("text", x = 116.5, y = 0.5, label = "region of positive\nselection identified\nin Winkler et al.")
```
(ref:IAA6-div-plot) **IAA6 diversity plot.** Nucleotide diversity of variant positions throughout the IAA6 coding sequence are plotted and colored according to the effect of the variant alleles at each position. The region of positive selection identified by Winkler et al. is highlighted.

### 2019-04-02 manuscript/index.Rmd
revisions done, back to coauthors

date: "`r Sys.Date()`"
abstract: "Thousands of sequenced genomes are now publicly available capturing a significant amount of natural variation within plant species; yet, much of this data remains inaccessible to researchers without significant bioinformatics experience. Here, we present a webtool called ViVa (Visualizing Variation) which aims to empower any researcher to take advantage of the amazing genetic resource collected in the *Arabidopsis thaliana* 1001 Genomes Project (http://1001genomes.org). ViVa facilitates data mining on the gene, gene family or gene network level. To test the utility and accessibility of ViVa, we assembled a team with a range of expertise within biology and bioinformatics to analyze the natural variation within the well-studied nuclear auxin signaling pathway. Our analysis has provided further confirmation of existing knowledge and has also helped generate new hypotheses regarding this well studied pathway. These results highlight how natural variation could be used to generate and test hypotheses about less studied gene families and networks, especially when paired with biochemical and genetic characterization. ViVa is also readily extensible to databases of interspecific genetic variation in plants as well as other organisms, such as the 3,000 Rice Genomes Project (http://snp-seek.irri.org/) and human genetic variation (https://www.ncbi.nlm.nih.gov/clinvar/).

@@

"

@@

#devtools::install_github("wrightrc/r1001genomes", ref = "auxin-natural-variation")
library(bookdown)
library(ape)
library(r1001genomes)
library(DT)
library(tidyverse)
library(magrittr)
library(RColorBrewer)
library(ggpmisc)
library(gginnards)
library(ggthemes)
library(ggseqlogo)
library(ggtree)
library(reticulate)
library(scales)
library(viridis)
library(ips)
library(knitr)
library(dplyr)
library(stringr)
library(treeio)

>>>>

date: "`r Sys.Date()`"
abstract: "Thousands of sequenced genomes are now publicly available capturing a significant amount of natural variation within plant species; yet, much of this data remains inaccessible to researchers without significant bioinformatics experience. Here, we present a webtool called ViVa (Visualizing Variation) which aims to empower any researcher to take advantage of the amazing genetic resource collected in the *Arabidopsis thaliana* 1001 Genomes Project (http://1001genomes.org). ViVa facilitates data mining on the gene, gene family or gene network level. To test the utility and accessibility of ViVa, we assembled a team with a range of expertise within biology and bioinformatics to analyze the natural variation within the well-studied nuclear auxin signaling pathway. Our analysis has provided further confirmation of existing knowledge and has also helped generate new hypotheses regarding this well studied pathway. These results highlight how natural variation could be used to generate and test hypotheses about less studied gene families and networks, especially when paired with biochemical and genetic characterization. ViVa is also readily extensible to databases of interspecific genetic variation in plants as well as other organisms, such as the 3,000 Rice Genomes Project (http://snp-seek.irri.org/) and human genetic variation (https://www.ncbi.nlm.nih.gov/clinvar/). "

@@

@@

#devtools::install_github("wrightrc/r1001genomes", ref = "auxin-natural-variation")
library(bookdown)
library(BALCONY)
library(ape)
library(r1001genomes)
library(DT)
library(tidyverse)
library(magrittr)
library(RColorBrewer)
library(ggpmisc)
library(gginnards)
library(ggthemes)
library(ggseqlogo)
library(ggtree)
library(reticulate)
library(scales)
library(viridis)
library(ips)
library(knitr)
library(dplyr)
library(stringr)
library(treeio)
